---
title: "Evanston Hypothesis Testing"
author: "Adi Tyagi"
date: "8/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading data and libraries
```{r}
library(tidyverse)
library(ggplot2)

#test hypotheses on evanston only data
fa_pm_evanston = read_csv("../Data/Modelling Data/fa_pm_evanston-model.csv")
fa_pm_evanston %>% head()
```

# Splitting the data
The top 5 most food insecure tracts are separate from the bottom 5 least food insecure tracts. This allocation is made based on 2018_food_insecurity_num. 


```{r}
#top 5 tracts
top_n = fa_pm_evanston %>% arrange(desc(food_insecurity_num_2018))%>% head(5) 
print(paste("Most food insecure tracts:",  top_n$local_census_tract))
```

```{r}
#bottom 5 tracts 
bottom_n = fa_pm_evanston %>% arrange(desc(food_insecurity_num_2018)) %>% tail(5)

print(paste("Least food insecure tracts: " ,bottom_n$local_census_tract))
```

# Distribution of number of food insecure persons in 2018
There are some tracts with immense food insecurity (800 persons); while others have ~200 food insecure persons. 
```{r, echo=FALSE}
fa_pm_evanston %>% ggplot() + geom_histogram(aes(food_insecurity_num_2018), binwidth = 40)

```

# Exploring differences in explanatory variables
## A simple analysis of difference

```{r}
var_list = c( 'median_household_income',
'median_age', 'social_vulnerability_index',
'median_home_loan_amount', 'median_home_value', 'prop_nonwhite',
'prop_nonenglish_speaking', 'proportion_hispanic',
'thiel_racial_segregation_index', 'proportion_bachelors_degree',
'computer_access', 'prop_families_poverty', 'num_jobs',
'num_housing_units', 'life_expectancy', 'avg_travel_time_to_work',
'prop_students_in_public_school', 'median_leverage_ratio',
'proportion_disabled', 'avg_household_size', 'prop_men')

for (var in var_list) {
  print(var)
  top_n_mean = mean(top_n[[var]])
  bottom_n_mean = mean(bottom_n[[var]])
  diff = top_n_mean - bottom_n_mean
  print(paste("top_n_mean:", top_n_mean))
  print(paste("bottom_n_mean:", bottom_n_mean))
  print(paste("diff:", top_n_mean - bottom_n_mean))
  print("------")
}
```

## Hypothesis testing

Let's test hypotheses on the following list of variables: 


- "median_household_income"        
- "median_age"                    
- "social_vulnerability_index"    
- "median_home_loan_amount"       
- "median_home_value"              
- "prop_nonwhite"                 
- "prop_nonenglish_speaking"       
- "proportion_hispanic"           
- "thiel_racial_segregation_index" 
- "proportion_bachelors_degree"   
- "computer_access"                
- "prop_families_poverty"         
- "num_jobs"                       
- "num_housing_units"             
- "life_expectancy"                
- "avg_travel_time_to_work"       
- "prop_students_in_public_school" 
- "median_leverage_ratio"         
- "proportion_disabled"            
- "avg_household_size"
- "prop_men"                    
- "local_census_tract"          
- "unemployment_change"


For each variable we will do a two-sided first. Then, if appropriate, we wil do a one-sided test. We will use a significance level of 10%. In some cases, we will use a significance level of 15%.  

### median household income
Let's begin with median_household_income
```{r}
t.test(top_n$median_household_income,bottom_n$median_household_income,
       alternative = "less")

```
Higher food insecurity tracts do indeed have a lower household income than do lower food insecurity tracts.

### median age
Let's now do median age. 
```{r}
t.test(top_n$median_age, bottom_n$median_age, alternative = "less")


```
Higher food insecurity tracts have a lower age. 

### social vulnerability index

```{r}
t.test(top_n$social_vulnerability_index, bottom_n$social_vulnerability_index, alternative = 'greater')

```

Food insecure tracts are more socially vulnerable. 

### median home loan amount
```{r}
t.test(top_n$median_home_loan_amount, bottom_n$median_home_loan_amount, alternative = 'less')

```

Food insecure tracts have a lower median home loan amount. 

### median home value
```{r}
t.test(top_n$median_home_value, bottom_n$median_home_value, alternative = 'less')
```

Food insecure tracts have lower median home value.

### prop nonwhite
```{r}
t.test(top_n$prop_nonwhite, bottom_n$prop_nonwhite, alternative = 'greater')
```

Food insecure tracts are more nonwhite in general.



### prop nonenglish speaking
```{r}
t.test(top_n$prop_nonenglish_speaking, bottom_n$prop_nonenglish_speaking, alternative = 'greater')
```

Food insecure tracts have a greater non english speaking proportion.

### median home value
```{r}
t.test(top_n$proportion_hispanic, bottom_n$proportion_hispanic, alternative = 'greater')
```

Food insecure tracts have a higher hispanic population.

### thiel racial segregation
```{r}
t.test(top_n$thiel_racial_segregation_index, bottom_n$thiel_racial_segregation_index, alternative = 'less')
```
Food insecure tracts are less racial segregated. 

### proportion bachelor's degree
```{r}
t.test(top_n$proportion_bachelors_degree, bottom_n$proportion_bachelors_degree, alternative = 'less')
```

Food insecure tracts have lower proportions with bachelor's degree.

### computer access
```{r}
t.test(top_n$computer_access, bottom_n$computer_access, alternative = 'less')
```
Food insecure tracts have lower computer access.

### prop families poverty
```{r}
t.test(top_n$prop_families_poverty, bottom_n$prop_families_poverty, alternative = 'greater')
```

Food secure tracts are similar to food insecure tracts when it comes to prop families in poverty.

### num jobs
```{r}
t.test(top_n$num_jobs, bottom_n$num_jobs, alternative = 'greater')
```
We could say that food insecure tracts have more jobs located within them than food secure tracts.

### num housing units
```{r}
t.test(top_n$num_housing_units, bottom_n$num_housing_units, alternative = 'greater')
```
Food insecure tracts are similar to food secure tracts when it comes to housing units.

### life expectancy
```{r}
t.test(top_n$life_expectancy, bottom_n$life_expectancy, alternative = 'less')
```
Food insecure tracts are similar to food secure tracts when it comes to  life expectancy.


### avg travel time to work
```{r}
t.test(top_n$avg_travel_time_to_work, bottom_n$avg_travel_time_to_work, alternative = 'less')
```

Food insecure tracts have lower avg travel times to work than food secure tracts.

### prop students in public school
```{r}
t.test(top_n$prop_students_in_public_school, bottom_n$prop_students_in_public_school, alternative = 'two.sided')
```
Food insecure and food secure tracts are similar when it comes to prop students in public school.

### median_leverage_ratio
```{r}
t.test(top_n$median_leverage_ratio, bottom_n$median_leverage_ratio, alternative = 'less')
```
Food secure vs. food insecure tracts are similar when it comes to leverage ratios. 


### prop disabled
```{r}
t.test(top_n$proportion_disabled, bottom_n$proportion_disabled, alternative = 'two.sided')
```
The food insecure tracts and the food secure tracts are similar in terms of prop disabled. 

### avg_household_size
```{r}
t.test(top_n$avg_household_size, bottom_n$avg_household_size, alternative = 'less')
```
Food insecure and food secure tracts are similar in terms of avg household size. 

### prop_men
```{r}
t.test(top_n$prop_men, bottom_n$prop_men, alternative = 'less')
```
Both food insecure and food secure tracts are similar when it comes to the proportion of men. 

# Takeaways

The following are the takeaways from the hypothesis testing. We used a significance level of 10%; for a few hypotheses we used a significance level of 15%. 

Compared to food secure tracts, food insecure tracts have/are:

- a lower household income
- a lower age. 
- more socially vulnerable 
- lower median home loan amount.
- lower median home value.
- more nonwhite in general.
- greater non english speaking proportion.
- higher hispanic population.
- less racial segregated. 
- lower proportions with bachelor's degree.
- lower computer access.
- lower avg travel times to work



Food insecure tracts and food secure tracts are similar in terms of: 

- prop families in poverty.
- num jobs
- num housing units
- life expectancy
- prop students in public school
- median leverage ratios
- prop disabled
- avg household size
- prop men